---
title: "PSET 4 - Problem 2"
author: "T"
date: "4/27/2020"
output: html_document
---

___WORK TO BE DONE___
_1) Finish analysis of comparing the two nests_
_2) Add Outputs and Tables to "Interpretations"_
_3) Debug issue of ~{ P_ij ≥ 1 }~_
_4) Dashboard_


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
heat_data <- read_csv("~/Documents/R/132/PSET 4/heating_data.csv")
source("~/Documents/Git/heat_demand/q2funs.R")

ggplot2::ggplot(select(heat_data, consumer_id, ic_gc)) +
  geom_point(aes(x = consumer_id, y = ic_gc)) +
  geom_hline(yintercept = 776.8, color = "#30c758", size = 2)
```

2a
What is the average installation and operating cost of each type of heating system?
What fraction of households choose each system type?
```{r}
library(tidyverse)
library(dplyr)

# Average installation and operation cost of each type of heating system
select(heat_data, contains("ic_"), contains("oc_"), X1) %>% 
  mutate_all(funs(average = mean(.))) %>%
  filter(X1 == 1) %>% 
  select(contains("average"))

# Fraction of household for each system
select(heat_data, choice) %>%
  group_by(choice) %>% 
  dplyr::summarise(count = length(choice)) %>% 
  mutate(fraction = round(count / sum(count), 2))
```

2b
Estimate a multinomial logit model of demand, including only installation cost and operating cost as product characteristics.
```{r}
b1 <- -.03
b2 <- -.01
betab <- c(b1, b2)
b <- optim(betab, fnb)
```
i.
Both coefficients have an negative sign. This is expected; a higher installment or operation cost will result in less
utility for a consumer.

ii.
73 cents. This is reasonable.
The installation cost is far greater than the operation cost. Adding money to the upfront cost is less appealing for two reasons. First, the installation cost is immediate and imposing. Necessary to have the heating system, the payment needs to be made straight away. Making this upfront cost larger makes the investment feel greater, from a short-term perspective. Second, increasing the cost of the larger, immediate component feels like a net-larger increase (from a short-run perspective). Additionally, if the lower cost (operation cost), which is paid over a year, is even lower it doesn't feel as impactful as, say, the installation cost decreasing.
For these reasons a 73 cent increase in upfront costs resulting from a one dollar decrease in operation costs seems reasonable.

iii.
```{r}
b1 <- -.03
b2 <- -.01
betab <- c(b1, b2)
fnb3(betab) # Calculates the average elasticities
```
Range of average elasticities for installation cost = (-31.24, -4.46).
Range of average elasticities for operation cost = (-4.44, -0.311).
On average, neither type of cost is inelastic, though operation cost is far less elastic than installation cost. Operation cost is closer to being, and for Gas Central actually is, inelastic.

2c
Multinomial logit model including product fixed effects, in addition to installation and operating costs.
```{r}
library(zeallot)
betas <- c(-.1, -.3, 0, 0, 0, 0)
c(b1, b2, b3, b4, b5, b6) %<-% betas
betac <- c(b1, b2, b3, b4, b5, b6)
c <- optim(betac, fnc)
```
Cost coefficients are not significantly different than without fixed effects. Ceteris paribus, the coefficients imply that relative to the Heat Pump heating system, customers prefer, in order: Electric Central, Electric Room, Gas Central, Gas Room.


2d
Modify the model in part 2c to include interactions between income and the two cost variables
```{r}
betas <- c(-.1, -.3, .1, .2, .15, .25, .02, .03)
c(b1, b2, b3, b4, b5, b6, b7, b8) %<-% betas
betad <- c(b1, b2, b3, b4, b5, b6, b7, b8)
d <- optim(betad, fnd)
```
$\beta_7$, installation cost * income, = -0.00021. $\beta_8$ operating cost * income, = 0.000017. $\beta_7$ is negative and indicates that sensitivity to the upfront, installation cost decreases as income increase. This makes sense. With higher income, upfront costs are less likely to be an issue. $\beta_8$ is positive and indicates that sensitivity to the longer-term, operation cost increases as income increases and vice versa. It is not intuitive to me that this is a reasonable value of $\beta_8is $. It seems that as income increases, aversity to higher costs, in both short- and long-term, should decrease. And vice versa.

One alternative interpretation of the coefficient on operating cost is that wealthier people are more conscious of long term spending. It doesn't seem unreasonable to assume that wealthier people have, on average, a more mindful approach to long term investments. Thus for someone with higher income, it wouldn't be strange that they are more sensitive to changes in operating costs than someone with lower income. This neither is a strong enough assumption nor one that has a sufficiently powerful effect to convinvce me that high income individuals are more sensitive to operating costs than low income. Both coefficients should be negative.


2e
Modify the model in part 2d to include interactions between income and the product fixed effects
```{r}
betas <- c(-.1, -.3, .1, .2, .15, .25, .02, .03, 0.01, 0.02, 0.01, 0.02)
c(b1, b2, b3, b4, b5, b6, b7, b8, b9, b10, b11, b12) %<-% betas
betae <- c(b1, b2, b3, b4, b5, b6, b7, b8, b9, b10, b11, b12)
e <- optim(betae, fne)
# b1 & b2: ic and oc
# b3-b6:
```
i. $\beta_7$ and $\beta_8$ are both negative. This implies that as income increases, sensitivity to both installation cost and operation cost decreases. This is intuitive: more income leads to less concern regarding prices.

ii. $\beta_9 = 0.167, \ \beta_{10} = 0.043, \ \beta_{11} = 0.017, \ \beta_{12} = 0.120$. These $\beta s$ represents the change in preference of Gas Central, Gas Room, Electric Central, and Electric Room respectively, over Heat Pump. Take Gas Central, for example. If the income of an individual increase by 1 unit ($10,000), then the observable utility of the individual of choosing Gas Central over Heat Pump increase by 0.167. The larger the coefficient of a product, the more that an individual with higher income prefers that heating system over the Heat Pump system, ceteris paribus. This may suggest that Gas Central and Electric Room are more expensive. It also may suggest that these are preferences only above a certain income threshold.

2f
Nested Logit. Nest by type of power (gas or electricity)
```{r}
betas <- c(-.1, -.3, .1, .2, .15, .25, .02, .03, 0.01, 0.02, 0.01, 0.02, 0.5, 0.8)
c(b1, b2, b3, b4, b5, b6, b7, b8, b9, b10, b11, b12, b13, b14) %<-% betas
betaf <- c(b1, b2, b3, b4, b5, b6, b7, b8, b9, b10, b11, b12, b13, b14)
nest1 <- optim(betaf, fnf)
# Yet to adjust the lambda parameters to inspect different outcomes.
# Before doing so I want to figure out how to increase the speed of that code
```

2fi
Nested Logit. Nest by type of system (central air or room)
```{r}
betas <- c(-.1, -.3, .1, .2, .15, .25, .02, .03, 0.01, 0.02, 0.01, 0.02, 0.5, 0.8)
c(b1, b2, b3, b4, b5, b6, b7, b8, b9, b10, b11, b12, b13, b14) %<-% betas
betafi <- c(b1, b2, b3, b4, b5, b6, b7, b8, b9, b10, b11, b12, b13, b14)
nest2 <- optim(betafi, fnfi)
```

2fii
Nested Logit Comparison
```{r}
NestComparison <-
  tibble(
    CoefficientOf = c("ic", "oc", "gcfixed", "grfixed", "ecfixed", "erfixed", 
            "ic*income", "oc*income", "gcfixed*income", "grfixed*income", 
            "ecfixed*income", "erfixed*income", "nest1weight", "nest2weight"),
    NestType1 = format(round(unlist(nest1$par), 6), scientific = FALSE),
    NestType2 = format(round(unlist(nest2$par), 6), scientific = FALSE),
    Description = coefficient_descriptions
  )
```
There is a meaningful difference between the two different choices of nesting. Nesting by central air or room systems as opposed to gas or electric systems caused changes in magnitude for many coefficients. Perhaps most importantly, we have some sign changes. Under the air vs room nesting, individuals now rank Gas Central as the preferred system, ceteris paribus, relative to Heat Pump. This is in contrast to gas vs electric nesting where Gas Central was the least preferred and actually yielded negative utility. Accompanying this change in sign is Electric Room becomes the least preferred under the air-room nesting whereas it was the 2nd ranked preference under the gas-electric nesting.

 ---- Finish Analysis
 




2g
```{r}
rebate <- heat_data %>% 
  mutate(ic_hp = .9*ic_hp)

betas <- c(-.1, -.3, .1, .2, .15, .25, .02, .03, 0.01, 0.02, 0.01, 0.02, 0.5, 0.8)
c(b1, b2, b3, b4, b5, b6, b7, b8, b9, b10, b11, b12, b13, b14) %<-% betas
betafi <- c(b1, b2, b3, b4, b5, b6, b7, b8, b9, b10, b11, b12, b13, b14)
betag <- c(b1, b2, b3, b4, b5, b6, b7, b8, b9, b10, b11, b12, b13, b14)
nest2 <- optim(betafi, fnfi)
rebatenest2 <- optim(betag, fng)



betas1 <- c(nest2$par)
betas2 <- c(rebatenest2$par)
beta_op_actual <- betas1
beta_op_rebate <- betas2


predictMarketShare(beta_op_actual, heat_data)
predictMarketShare(beta_op_rebate, rebate)
```
With the rebate, the model predicts a 9% market share for heat pumps (assuming that consumers choose the product with the greatest probability). This is significantly greater than the 2% market share for heat pumps without the 10% subsidy!




__Comparison__
Think about the models on their own merit without actually looking at the results.


Less Important - but likely useful for dashboard purposes and good to make general. Easy to generalize.
The issue that I am facing is having the tibble full of doubles / numeric in order to plot above and below zero, with zero as the baseline in the middle of the plot.
Attempt to make a divergent bar plot to visualize the different coefficients
```{r}
library(tibble)

LogitComparison <-
  tibble(
    CoefficientOf = c("ic", "oc", "gcfixed", "grfixed", "ecfixed", "erfixed", 
            "ic_income", "oc_income", "gcfixed_income", "grfixed_income", 
            "ecfixed_income", "erfixed_income", "nest1weight", "nest2weight"),
    b = c(round(b$par,5), as.numeric(rep("0", 12))),
    c = c(round(c$par,5), as.numeric(rep("0", 8))),
    d = c(as.numeric(rep("0", 6)), round(d$par,6)),
    e = c(round(e$par,6), as.numeric(rep("0", 2))),
    NestType1 = format(round(unlist(nest1$par), 6), scientific = FALSE),
    NestType2 = format(round(unlist(nest2$par), 6), scientific = FALSE),
    Description = coefficient_descriptions
  )
View(LogitComparison)


coeff_comp_plot <- function() {
  
}
plot <- filter(LogitComparison, CoefficientOf == "gcfixed") %>% 
  gather(key = Model, gcfixed, b:NestType2) %>% 
  mutate(gcfixed_Coefficient = gcfixed)

ggplot(plot, aes(x = Model, y = gcfixed_Coefficient)) + 
  geom_bar(stat = "identity")
  coord_flip()
  labs(title = paste(plot$Description, "for each model used"))


tibble(
  x = c("b", "c", "d", "e", "f"),
  y = c(5, -5, -2, -.2, 0)
) %>% 
  ggplot(aes(x = x, y = y)) +
  geom_bar(stat = "identity", width = .5)
```
```
















